$p$-adic precision

### 2014-02-19 padicprec.tex
Working on Section 3.1

\newcommand{\noopsort}[1]{}

@@

\newcommand{\lb}{\ensuremath{\llbracket}}
\newcommand{\rb}{\ensuremath{\rrbracket}}
\newcommand{\lp}{(\!(}
\newcommand{\rp}{)\!)}

@@

If $E$ is a $K$-vector space, possibly of infinite dimension, then an
\emph{ultrametric norm} on $E$ is a map $\Vert\cdot\Vert : E \to \R^+$ satisfying:
\begin{enumerate}[(i)]
\item $\Vert x\Vert = 0$ if and only if $x = 0$;
\item $\Vert \lambda x\Vert = |\lambda| \cdot \Vert x\Vert$;
\item $\Vert x+y\Vert \leq \max(\Vert x\Vert, \Vert y\Vert)$.
\end{enumerate}
A \emph{$K$-Banach space} is a complete normed $K$-vector space.
A \emph{lattice} in a $K$-Banach space $E$ is an open bounded sub-$\OK$-module of $E$.

@@

On the other hand, one can prove that the SOMOS 4 sequence 
exhibits the \emph{Laurent phenomenon} \cite{fomin-zelevinsky:02a}. This means that, for all integer 
$n$, there exists a polynomial $P_n$ in $\Z[X^{\pm 1}, Y^{\pm 1}, Z^{\pm 
1}, T^{\pm 1}]$ such that $u_n = P_n(u_0, u_1, u_2, u_3)$.
From the latter formula, it follows directly that if $u_0$, $u_1$,
$u_2$ and $u_3$ are known up to precision $O(p^N)$ then all $u_n$'s
are also known with the same precision. Thus, the term $v_n$ that 
appears in \eqref{eq:SOMOS} does not reflect an intrinsic loss of
precision but some numerical instability related to the algorithm used to compute $u_n$.

\begin{rmk}
From the above discussion, one can easily derive a numerically stable 
algorithm that computes the SOMOS 4 sequence: 
\begin{enumerate}
\item compute the Laurent polynomials $P_n$ using the induction formula (and working in 
the ring $\Z[X^{\pm 1}, Y^{\pm 1}, Z^{\pm 1}, T^{\pm 1}]$)
\item evaluate $P_n$ at the point $(u_0, u_1, u_2, u_3)$.
\end{itemize}
However, computing the $P_n$'s is very time-consuming since it requires 
division in a polynomial ring with $4$ variables and the
size of the coefficients of $P_n$ explodes as $n$ grows.

@@

@@

The theory of $p$-adic precision we want to develop is mainly based on a 
single lemma (Lemma \ref{lem:main}) of $p$-adic analysis. The aim of 
this Section is to state and prove this lemma. Consequences will be 
discussed after in Section \ref{sec:tracking}.

From now, we generalize a bit the setting and work over a general 
ultrametric complete field (see definition below). This includes 
$p$-adic fields and Laurent series fields.

\subsection{Ultrametric analysis}
\label{ssec:ultrametric}

Similarly to the real case, any finite dimensional normed $K$-vector 
space is complete and all norms over such a space are equivalent 
(\emph{i.e.} they define the same topology). Many usual Theorems 
concerning Banach spaces (\emph{e.g.} open mapping Theorem, closed graph 
Theorem) remain true in the ultrametric setting.

\begin{deftn} \label{deftn : diff}
Let $V, W$ be two normed $K$-vector spaces, let $U \subset V$ be an open 
subset of $V$ and let $f : U \rightarrow W$ be a map. Then $f$ is called 
\emph{differentiable} at $v_0 \in U$ if there exists a 
continuous linear map $f'(v_0) : U \rightarrow W$ such that for any 
$\varepsilon >0$, there exists an open neighborhood $U_\varepsilon 
\subset U$ containing $v_0$ with:
\[ 
\forall v, w \in U_\varepsilon, \quad
\Vert f(v)-f(w)-f'(v_0) \cdot \left( v-w \right) \Vert 
\leq \varepsilon \Vert v-w \Vert. 
\]
The linear map $f'(v_0)$ is called the \emph{differential} of $f$ at $v_0$.
\end{deftn}

@@

%The following Lemma shows that the notion of lattices behaves rather 
%well under linear\footnote{We shall see in the sequel (\emph{cf} Lemma 
%\ref{lem:main}) that, under some extra hypothesis, the linearity 
%assumption can be drastically relaxed.} transformations.

%\begin{lem}
%\label{lem:morlat}
%Let $E$ and $F$ be two $K$-Banach and $f : 
%E \to F$ be a continuous \emph{surjective} linear map.
%If $H$ is a lattice in $E$, then $f(H)$ is a lattice in $F$.
%\end{lem}

%\begin{proof}
%It is clear that $f(H)$ is a module over $B_K(1)$ which is bounded.
%The fact that it is open is a direct consequence of the open mapping
%Theorem.
%\end{proof}

%To each lattice $H$ in $E$, one can attach a norm $\Vert \cdot \Vert_H$ 
%on $E$ defined as follows. We pick $x \in E$, $x \neq 0$. We introduce
%the set $I_x$ consisting of scalars $\lambda \in K$ such that $\lambda x 
%\in H$. It follows from the fact that $H$ is open and bounded in $E$
%that $I_x$ is nonempty and bounded in $K$. Therefore, the image $|I_x|$ 
%of $I_x$ under the absolute value map is bounded and contains a positive
%element. We can then define:
%$$\Vert x \Vert_H = (\sup |I_x|)^{-1}.$$
%It is an exercise to check that $\Vert \cdot \Vert_H$ is a norm over 
%$E$.

%\begin{lem} \label{lem:reductionball}
%\begin{enumerate}
%\item The norms $\Vert \cdot \Vert_H$ and $\Vert \cdot \Vert$ are 
%equivalent.
%\item Let $B^-_{E,H}(1)$ (resp. by $B_{E,H}(1)$) the set of elements $x 
%\in E$ such that $\Vert x \Vert_H < 1$ (resp. $\Vert x \Vert_H \leq 1$).
%Then:
%$$B^-_{E,H}(1) \subset H \subset B_{E,H}(1)$$
%and the last inclusion is an equality if the absolute value on $K$ is
%discrete.
%\end{enumerate}
%\end{lem}

%\begin{proof}
%Left to the reader.
%\end{proof}

\subsubsection*{Lattices}

@@

\begin{deftn}
\label{def:firstorder}
Let $E$ and $F$ be two $K$-Banach and $f : U \rightarrow F$ be a 
function defined on an open subset $U$ of $E$. Let also $v_0$ be a 
point in $U$.
A lattice $H$ in $E$ is called a \emph{first order lattice} for $f$ at 
$v_0$ if the equality
\begin{equation}
\label{eq:firstorder}
f(v_0 + H) = f(v_0) + f'(v_0) (H)
\end{equation}
holds.
\end{deftn}

We emphasize that we require an equality in Eq.~\eqref{eq:firstorder}, 
and not just an inclusion! It will be very important for the 
applications we have in mind.
Here comes our main Lemma.

\begin{lem} \label{lem:main}
Let $E$ and $F$ be two $K$-Banach and $f : U 
\rightarrow F$ be a function defined on an open subset $U$ of $E$.
We assume that $f$ is differentiable at some point $v_0 \subset 
U$ and that the differential $f'(v_0)$ is surjective.

@@

\begin{proof}
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, by 
the open mapping Theorem, there exists $C>0$ such that $B_F(1) \subset 
C  f'(0)(B_E(1))$. Let $\varepsilon>0$ be such that 
$\varepsilon C < \rho$. Let $U_\varepsilon \subset E$ given by Definition 
\ref{deftn : diff} and the fact that $f$ is differentiable at $0$. We 
may assume $U_\varepsilon = B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$. Let $B^-_E(\rho r) \subset H \subset B_E(r).$ Let us 
show that $f$ maps surjectively $H$ onto $f'(0) (H)$. We first prove 
that $f(H) \subset f'(0) (H)$. Let $x \in H$. By differentiability at 
$0$, $\Vert f(x)-f'(0)(x) \Vert \leq \varepsilon \Vert x \Vert $. 
Therefore, if $y=f(x)-f'(0)(x)$, then $\Vert y \Vert \leq 
\varepsilon r$. By definition of $C$, we have $B_F(\varepsilon r) 
\subset \varepsilon C f'(0) (B_E(r))$. Thus there exists $x_1 \in 
\varepsilon C \cdot B_E(r)$ such that $f'(0) (x_1) =y$. Since 
$\varepsilon C < \rho$, we get $x_1 \in B^-_E(\rho r) \subset H$ and 
then $f(x)= f'(0) (x-x_1) \in f'(0) (H)$.

We shall now prove surjectivity. Let $y \in f'(0) (H)$. Let $x_0 \in H$ 
be such that $y = f'(0) (x_0)$. We define inductively two sequences 
$(x_n)$ and $(z_n)$ as follows:
\begin{itemize}
\item $z_n$ is an element of $E$ satisfying $f'(0)(z_n) = y - 
f(x_n)$ and $\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$ (such 
an element exists by definition of $C$), and
\item $x_{n+1}=x_n+z_n$.
\end{itemize}
For convenience, let us also define $x_{-1} = 0$ and $z_{-1}=x_0.$ We claim that the 
sequences $(x_n)$ and $(z_n)$ are well defined and take their values in 
$H$. To establish this, it is enough to prove that, assuming that 
$x_{n-1}$ and $x_n$ belongs to $H$, so do $z_n$ and $x_{n+1}$. Noticing
that 
\begin{equation}
\label{eq:mainlemma}
y - f(x_n) 
= f(x_{n-1}) + f'(0)(z_{n-1}) - f(x_n)
= f(x_{n-1}) - f(x_n) - f'(0)(x_{n-1} - x_n)
\end{equation}
we deduce using differentiability that
$\Vert y - f(x_n) \Vert \leq \varepsilon \cdot \Vert x_n - x_{n-1}
\Vert$.
Since we are assuming that $x_{n-1}$ and $x_n$ lie in $H \subset
B_E(r)$, we find $\Vert y - f(x_n) \Vert \leq \varepsilon r$. Thus
$\Vert z_n \Vert \leq C \cdot \varepsilon r < \rho r$ and then
$z_n \in H$. From the relation $x_{n+1} = x_n + z_n$, we finally
deduce $x_{n+1} \in H$.

Now using again Eq.~\eqref{eq:mainlemma} and differentiability at $0$,
we get, for all positive integer $n$,
$$\Vert y  - f(x_n) \Vert 
\leq \varepsilon \cdot \Vert z_{n-1} \Vert \leq \varepsilon C \cdot \Vert 
y - f(x_{n-1}) \Vert,$$
from what we deduce that $\Vert y - f(x_n) \Vert = O(a^n)$ and 
$\Vert z_n \Vert = O(a^n)$ with $a = \varepsilon C < \rho \leq 1$. 
This shows that $(x_n)$ is a Cauchy sequence and, since $E$ is complete, 
it converges. Because $H$ is closed, if $x$ denotes the limit of 
$(x_n)$, we have $x \in H$. Moreover, as a consequence of the 
differentiability, $f$ is continuous over $U_\varepsilon$. Therefore, 
since $x \in H \subset U_\varepsilon$, we get $y=f(x)$ and we are done.
\end{proof}

Let us conclude this paragraph by a remark on the assumption made in 
Lemma \ref{lem:main} about the surjectivity of $f'(v_0)$. First let us 
emphasize that this hypothesis is definitely necessary. Indeed, it would 
implies otherwise that the image of $f$ is locally contained in a proper 
sub vector space around each point where the differential of $f$ is not 
surjective, which is certainly not true! Nevertheless, one can get 
around this issue in many situations. Indeed, consider the situation of 
Lemma \ref{lem:main} without assuming that $f'(v_0)$ is surjective. Now, 
we pick a closed sub vector space $K$ of $F$ such that $K + f'(v_0)(E) = 
F$. Then, denoting by $\pr_K$ the canonical projection of $F$ onto 
$F/K$, the composite $\pr_K \circ f$ is certainly differentiable at 
$v_0$ at its differential is surjective. Hence Lemma \ref{lem:main}
applies and shows that
\begin{equation}
\label{eq:notsurjective}
f(v_0 + H) \subset f(v_0) + f'(v_0)(H) + K
\end{equation}
for all lattices $H$ satisfying a certain condition. Now we can take
the intersection of the right hand side of \eqref{eq:notsurjective} 
over all possible $K$ and hope that we get this way a sharp affine upper 
bound of $f(v_0 + H)$.

@@

Before discussing applications of Lemma \ref{lem:main} to ultrametric 
precision, we would like to describe a little bit more restrictive 
setting where the constant $\delta$ appearing in this Lemma can be 
computed explicitly.

Indeed, when $f$ is a locally analytic function, we prove that we can recover $\delta$ from a Newton polygon defined with the Taylor series expansion of $f.$

We refer to Schneider \cite{schneider:11a}, \S 6, for a definition of a locally analytic function, and we extend it to any complete $K$ vector space :

\begin{deftn} \label{deftn:analyticNP}
Let $E$ and $F$ be two complete ultrametric $K$-Banach spaces. Let $U \subset E$ be an open set.
A function $f \: : \: U \rightarrow F$ is said to be locally analytic if for any $x \in U,$ there exists some open subset $U_x$ of $E$ and some $n$-linear mappings $L_n \: : \: E^n \rightarrow F$ for any $n \geq 1,$ such that : for any $y \in U_x,$ the sum $\sum_{n \geq 1} L_n(y,\dots,y)$ converges in F, and $f(x+y)=f(x)+\sum_{n \geq 1} L_n(y,\dots,y).$
\end{deftn}

@@

\begin{deftn} \label{deftn :  serie and norm}
Let $f(x)= \sum_{n \geq 0} f_n (x)$ be a serie over mappings $E \rightarrow F,$ where $E$ and $F$ are complete normed $\overline{K}$-vector space. We assume that for any $\lambda  \in \overline{K},$ $x \in E,$ $n \in \mathbb{N},$ $\vert f_n(\lambda x) \vert = \vert \lambda \vert \vert f_n (x) \vert.$ 
We define $\vert f_n \vert = \sup_{x \in B_E(0,1)} \vert f_n(x) \vert,$ and we assume that all the $\vert f_n \vert $ are finite. 
We also define, for any $v \in \mathbb{R}$ such that this $\sup$ is finite : \[\Lambda(f) (v) = \log \sup_{x \in B_E(0,e^v)} \vert f(x) \vert,\]
and $\Lambda(f) (v) =+\infty$ otherwise.
We define  $NP(f)$ to be the Newton polygon of $f$, that is to say, the lower convex hull in $\mathbb{R}^2$ of the $(k, - \log \vert f_k \vert ),$ $k\geq 0.$ 
\end{deftn}

@@

\begin{prop} \label{prop : conjugate Newton}
We define $v_\infty \in \mathbb{R} \cup \left\lbrace \infty \right\rbrace$ to be the limit of the slopes of $NP(f).$
We denote by $NP(f)^*$ the convex conjugate of $NP(f).$ $NP(f)^*$ is defined over $\left] -\infty,v_{\infty} \right[$ (at least)
\end{prop}

Then, $\Lambda(f)$ and $NP(f)$ are connected :

\begin{prop} \label{prop : connexion sup Newton}
For any $v \in \left]-\infty, v_\infty \right[,$ \[ \Lambda (f) (v) = NP(f)^*(v).\]
\end{prop}
\begin{proof}
Let $v \in \left] -\infty,v_{\infty} \right[,$ and let us assume that $v$ is not a slope of $NP(f).$
Let $u = NP(f)^*(v).$ Then the intersection of the line defined by $y=vx-u$ and $NP(f)$ is an extreme point of $NP(f),$ say $(n, - \log ( \vert f_n \vert ))$ for some $n \geq 0,$ and $u = vn + \log \vert f_n \vert.$

$v$ is not a slope of $NP(f)$ and by convexity of $NP(f)$, its difference quotients at $n$ are increasing, therefore, there exists some $c>0$ such that for any $m\geq 0,$ $vm-u \leq - \log \vert f_m \vert - c \vert n-m \vert.$ Since $vm-u=vm-vn-log \vert f_n \vert,$ we get $-vn-\log \vert f_n \vert + c \vert n-m \vert  \leq -vm-\log \vert f_m \vert .$

@@

As a consequence, $\lim_i \vert f(x_i) \vert = \vert f_n \vert e^{vn},$ and  $\Lambda (f)(v) = \log ( \vert f_n \vert e^{vn}) = u= NP(f)^*(v).$

Therefore, for any $v \in \left] -\infty,v_{\infty} \right[,$ with $v$ not a slope of $NP(f),$ we get $\Lambda (f)(v) =NP(f)^*(v).$
Since $\Lambda (f)$ is increasing, $NP(f)^*$ is continuous (since it is convex), and $NP(f)$ has only finitely many slopes smaller than $v_\infty-\epsilon$, for any $\epsilon >0$, we deduce that both functions coincide on $v \in \left] -\infty,v_{\infty} \right[,$
\end{proof}

@@

The following proposition will reveal crucial in order to minorate or majorate $NP(f),$  $NP(f)^*$ or $\Lambda(f),$ and is classical in convex analysis.

@@

If we denote by $NP((M_{n+1})_{n \geq 1})$ the lower convex hull of the $(n, -\log M_{n+1})$ for $n \geq 1$, then $\delta$ such that $-\log \varepsilon=NP((M_n)_{n \geq 2})^*(r)$ is enough to define first order lattices as in lemma  \ref{lem:main}.
\end{prop}
\begin{proof}
Same idea as proposition \ref{prop : maj Lambda}. One can define $g_n$ with $g_n (x) =\frac{f_n(x)}{\vert x \vert}$ and $g_n(0)=0$ and work with $\Lambda ( \sum g_n).$
\end{proof}

@@

\begin{prop} \label{prop : locanalyticfinitedim}
If $f$ is a locally analytic map $K^n \rightarrow K^m$, with $K$ a $p$-adic field, if $C,$ $\rho$ and $\varepsilon$ are as above. It is enough to work with the Newton polygon $NP_F$ of $F(t) = \sum_{k\geq 0}^{+\infty} \interleave f^{(k+1)}(0) \interleave  t^k \in \mathbb{R}[[t]]$ : let $r$ be such that $NP_F^* \left( \log_p r - \frac{1}{p-1} \right) \leq \log_p \varepsilon.$ Then $r$ is enough for lemma \ref{lem:main}. 
\end{prop}
\begin{proof}
Same as before, with $L_n=\frac{f^{(n)}(0)}{n!}$ and Legendre formula. The result follows \textit{mutatis mutandis}.
\end{proof}

@@

In what follows, the letter $K$ always refers to an ultrametric field 
(\emph{cf} \S \ref{ssec:ultrametric}). The absolute value over $K$ is 
denoted by $|\cdot|$.

>>>>

\newcommand{\noopsort}[1]{}
\DeclareMathOperator{\NP}{NP}

@@

\newcommand{\lb}{\ensuremath{\llbracket}}
\newcommand{\rb}{\ensuremath{\rrbracket}}
\newcommand{\lp}{(\!(}
\newcommand{\rp}{)\!)}
\newcommand{\col}{\: : \:}

@@

If $E$ is a $K$-vector space, possibly of infinite dimension, then an
\emph{ultrametric norm} on $E$ is a map $\Vert\cdot\Vert : E \to \R^+$ satisfying:
\begin{enumerate}[(i)]
\item $\Vert x\Vert = 0$ if and only if $x = 0$;
\item $\Vert \lambda x\Vert = |\lambda| \cdot \Vert x\Vert$;
\item $\Vert x+y\Vert \leq \max(\Vert x\Vert, \Vert y\Vert)$.
\end{enumerate}
A \emph{$K$-Banach space} is a complete normed $K$-vector space.  Note that any finite-dimensional
normed $K$-vector space is automatically complete and all norms over such a space are equivalent.
A \emph{lattice} in a $K$-Banach space $E$ is an open bounded sub-$\OK$-module of $E$.

@@

On the other hand, one can prove that the SOMOS 4 sequence 
exhibits the \emph{Laurent phenomenon} \cite{fomin-zelevinsky:02a}: for all integer 
$n$, there exists a polynomial $P_n$ in $\Z[X^{\pm 1}, Y^{\pm 1}, Z^{\pm 
1}, T^{\pm 1}]$ such that $u_n = P_n(u_0, u_1, u_2, u_3)$.
From the latter formula, it follows directly that if $u_0$, $u_1$,
$u_2$ and $u_3$ are known up to precision $O(p^N)$ then all $u_n$'s
are also known with the same precision. Thus, the term $v_n$ that 
appears in \eqref{eq:SOMOS} does not reflect an intrinsic loss of
precision but some numerical instability related to the algorithm used to compute $u_n$.

\begin{rmk}
From the above discussion, one can easily derive a numerically stable 
algorithm that computes the SOMOS 4 sequence: 
\begin{enumerate}
\item compute the Laurent polynomials $P_n$ using the recurrence in 
the ring $\Z[X^{\pm 1}, Y^{\pm 1}, Z^{\pm 1}, T^{\pm 1}]$
\item evaluate $P_n$ at the point $(u_0, u_1, u_2, u_3)$.
\end{enumerate}
However, computing the $P_n$'s is very time-consuming since it requires 
division in a polynomial ring with $4$ variables and the
size of the coefficients of $P_n$ explodes as $n$ grows.

@@

% Frame better

@@

Our theory of $p$-adic precision rests upon a 
lemma in $p$-adic analysis: Lemma \ref{lem:main}.  In this section
we develop the theory surrounding this result, and proceed to practical
consequences in Section \ref{sec:tracking}.

Our goal in this section is to relate the image of a lattice under a differentiable map
to its image under the derivative.

\begin{deftn} \label{deftn : diff}
Let $V, W$ be two normed $K$-vector spaces, let $U \subset V$ be an open 
subset of $V$ and let $f : U \rightarrow W$ be a map. Then $f$ is called 
\emph{differentiable} at $v_0 \in U$ if there exists a 
continuous linear map $f'(v_0) : U \rightarrow W$ such that for any 
$\varepsilon >0$, there exists an open neighborhood $U_\varepsilon 
\subset U$ containing $v_0$ with
\[ 
\Vert f(v)-f(w)-f'(v_0) \cdot \left( v-w \right) \Vert 
\leq \varepsilon \Vert v-w \Vert. 
\]
for all $v, w \in U_\varepsilon$.  The linear map $f'(v_0)$ is called the \emph{differential} of $f$ at $v_0$.
\end{deftn}

@@

@@

\begin{deftn}
\label{def:firstorder}
Let $E$ and $F$ be two $K$-Banach spaces, $f : U \rightarrow F$ be a 
function defined on an open subset $U$ of $E$ and $v_0$ be a 
point in $U$.
A lattice $H$ in $E$ is called a \emph{first order lattice} for $f$ at 
$v_0$ if the equality
\begin{equation}
\label{eq:firstorder}
f(v_0 + H) = f(v_0) + f'(v_0) (H)
\end{equation}
holds.
\end{deftn}

We emphasize that we require an equality in Eq.~\eqref{eq:firstorder}, 
and not just an inclusion! 
With this definition in hand, we are able to state our main lemma.

\begin{lem} \label{lem:main}
Let $E$ and $F$ be two $K$-Banach spaces and $f : U 
\rightarrow F$ be a function defined on an open subset $U$ of $E$.
We assume that $f$ is differentiable at some point $v_0 \in 
U$ and that the differential $f'(v_0)$ is surjective.

@@

\begin{proof} % Call it the Banach-Schauder theorem?
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, the open mapping theorem provides a $C>0$ such that 
\[
B_F(1) \subset C  f'(0)(B_E(1)).
\]
Let $\varepsilon>0$ be such that 
$\varepsilon C < \rho$, and choose $U_\varepsilon \subset E$ as in Definition 
\ref{deftn : diff}.  We may assume $U_\varepsilon = B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$ and suppose $H$ is a lattice with $B^-_E(\rho r) \subset H \subset B_E(r).$ 
We seek to show that $f$ maps $H$ surjectively onto $f'(0) (H)$. We first prove 
that $f(H) \subset f'(0) (H)$. Suppose $x \in H$; we first show that $f(x) \in f'(0)(H)$.  By differentiability at 
$0$, $\Vert f(x)-f'(0)(x) \Vert \leq \varepsilon \Vert x \Vert $. 
Setting $y=f(x)-f'(0)(x)$, we have $\Vert y \Vert \leq 
\varepsilon r$. The definition of $C$ implies that $B_F(\varepsilon r) 
\subset \varepsilon C f'(0) (B_E(r))$. Thus there exists $x' \in 
\varepsilon C \cdot B_E(r)$ such that $f'(0) (x') =y$. Since 
$\varepsilon C < \rho$, we get $x' \in B^-_E(\rho r) \subset H$ and 
then $f(x)= f'(0) (x-x') \in f'(0) (H)$.

We shall now prove surjectivity. Let $y \in f'(0) (H)$. Let $x_0 \in H$ 
be such that $y = f'(0) (x_0)$. We inductively define two sequences 
$(x_n)$ and $(z_n)$ as follows:
\begin{itemize}
\item $z_n$ is an element of $E$ satisfying $f'(0)(z_n) = y - 
f(x_n)$ and $\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$ (such 
an element exists by definition of $C$), and
\item $x_{n+1}=x_n+z_n$.
\end{itemize}
For convenience, let us also define $x_{-1} = 0$ and $z_{-1}=x_0.$ We claim that the 
sequences $(x_n)$ and $(z_n)$ are well defined and take their values in 
$H$. We do so by induction, assuming that $x_{n-1}$ and $x_n$ belong to $H$
and showing that $z_n$ and $x_{n+1}$ do as well. Noticing
that 
\begin{equation}
\label{eq:mainlemma}
\begin{aligned}
y - f(x_n) &= f(x_{n-1}) + f'(0)(z_{n-1}) - f(x_n) \\
&= f(x_{n-1}) - f(x_n) - f'(0)(x_{n-1} - x_n)
\end{aligned}
\end{equation}
we deduce using differentiability that
$\Vert y - f(x_n) \Vert \leq \varepsilon \cdot \Vert x_n - x_{n-1}
\Vert$.
Since we are assuming that $x_{n-1}$ and $x_n$ lie in $H \subset
B_E(r)$, we find $\Vert y - f(x_n) \Vert \leq \varepsilon r$. Thus
$\Vert z_n \Vert \leq C \cdot \varepsilon r < \rho r$ and then
$z_n \in H$. From the relation $x_{n+1} = x_n + z_n$, we finally
deduce $x_{n+1} \in H$.

Using Eq.~\eqref{eq:mainlemma} and differentiability at $0$ once more,
we get
$$\Vert y  - f(x_n) \Vert 
\leq \varepsilon \cdot \Vert z_{n-1} \Vert \leq \varepsilon C \cdot \Vert 
y - f(x_{n-1}) \Vert,$$
for all positive integer $n$.  We therefore deduce that $\Vert y - f(x_n) \Vert = O(a^n)$ and 
$\Vert z_n \Vert = O(a^n)$ with $a = \varepsilon C < \rho \leq 1$. 
These conditions show that $(x_n)$ is a Cauchy sequence, which converges since $E$ is complete.
Write $x$ for the limit of the $x_n$; we have $x \in H$ because $H$ is closed.
Moreover, $f$ is continuous on $H \subseteq U_\varepsilon$ since it is differentiable and thus $y=f(x)$.
\end{proof}

We end this section with a remark on the surjectivity of $f'(v_0)$ assumed in
Lemma \ref{lem:main}. First, let us 
emphasize that this hypothesis is definitely necessary. Indeed, the lemma would 
otherwise imply that the image of $f$ is locally contained in a proper 
sub-vector-space around each point where the differential of $f$ is not 
surjective, which is certainly not true! Nevertheless, one can use lemma \ref{lem:main}
to prove a weaker result in the context that $f'(v_0)$ is not surjective.
To do so, choose a closed sub-vector-space $W$ of $F$ such that $W + f'(v_0)(E) = 
F$. Denoting by $\pr_W$ the canonical projection of $F$ onto 
$F/W$, the composite $\pr_W \circ f$ is differentiable at 
$v_0$ with surjective differential. For a given lattice $H$, there will be various choices of $W$ to which Lemma \ref{lem:main}
applies.  For each such $W$,
\begin{equation}
\label{eq:notsurjective}
f(v_0 + H) \subset f(v_0) + f'(v_0)(H) + W;
\end{equation}
taking the intersection of the right hand side
over all possible $W$ yields an upper bound on $f(v_0 + H)$.

@@

Before discussing applications of Lemma \ref{lem:main} to ultrametric 
precision, we discuss the case that $f$ is locally analytic.  In this context
we prove that we can recover the constant $\delta$ appearing in the lemma
from the Newton polygon determined by the Taylor series expansion of $f$.

Schneider \cite{schneider:11a}*{\S 6} defines locally analytic functions on finite dimensional vector spaces.
His definition extends easily to $K$-Banach spaces:

\begin{deftn} \label{deftn:analyticNP}
Let $E$ and $F$ be $K$-Banach spaces, and let $U \subset E$ be an open set.
A function $f \col U \rightarrow F$ is said to be \emph{locally analytic} if for any $x \in U,$ there exists some open subset $U_x$ of $E$ and some $n$-linear mappings $L_n \: : \: E^n \rightarrow F$ for any $n \geq 1,$ such that : for any $y \in U_x,$ the sum $\sum_{n \geq 1} L_n(y,\dots,y)$ converges in F, and $f(x+y)=f(x)+\sum_{n \geq 1} L_n(y,\dots,y).$
\end{deftn}

@@

\begin{deftn} \label{deftn :  serie and norm}
Let $f(x)= \sum_{n \geq 0} f_n (x)$ be a serie over mappings $E \rightarrow F,$ where $E$ and $F$ are complete normed $\overline{K}$-vector space. We assume that for any $\lambda  \in \overline{K},$ $x \in E,$ $n \in \mathbb{N},$ $\vert f_n(\lambda x) \vert = \vert \lambda \vert \vert f_n (x) \vert.$ 
We define $\vert f_n \vert = \sup_{x \in B_E(0,1)} \vert f_n(x) \vert,$ and we assume that all the $\vert f_n \vert $ are finite. 
We also define, for any $v \in \mathbb{R}$ such that this $\sup$ is finite : \[\Lambda(f) (v) = \log \sup_{x \in B_E(0,e^v)} \vert f(x) \vert,\]
and $\Lambda(f) (v) =+\infty$ otherwise.
We define  $\NP(f)$ to be the Newton polygon of $f$, that is to say, the lower convex hull in $\mathbb{R}^2$ of the $(k, - \log \vert f_k \vert ),$ $k\geq 0.$ 
\end{deftn}

@@

\begin{prop} \label{prop : conjugate Newton}
We define $v_\infty \in \mathbb{R} \cup \left\lbrace \infty \right\rbrace$ to be the limit of the slopes of $\NP(f).$
We denote by $\NP(f)^*$ the convex conjugate of $\NP(f).$ $\NP(f)^*$ is defined over $\left] -\infty,v_{\infty} \right[$ (at least)
\end{prop}

Then, $\Lambda(f)$ and $\NP(f)$ are connected :

\begin{prop} \label{prop : connexion sup Newton}
For any $v \in \left]-\infty, v_\infty \right[,$ \[ \Lambda (f) (v) = \NP(f)^*(v).\]
\end{prop}
\begin{proof}
Let $v \in \left] -\infty,v_{\infty} \right[,$ and let us assume that $v$ is not a slope of $\NP(f).$
Let $u = \NP(f)^*(v).$ Then the intersection of the line defined by $y=vx-u$ and $\NP(f)$ is an extreme point of $\NP(f),$ say $(n, - \log ( \vert f_n \vert ))$ for some $n \geq 0,$ and $u = vn + \log \vert f_n \vert.$

$v$ is not a slope of $\NP(f)$ and by convexity of $\NP(f)$, its difference quotients at $n$ are increasing, therefore, there exists some $c>0$ such that for any $m\geq 0,$ $vm-u \leq - \log \vert f_m \vert - c \vert n-m \vert.$ Since $vm-u=vm-vn-log \vert f_n \vert,$ we get $-vn-\log \vert f_n \vert + c \vert n-m \vert  \leq -vm-\log \vert f_m \vert .$

@@

As a consequence, $\lim_i \vert f(x_i) \vert = \vert f_n \vert e^{vn},$ and  $\Lambda (f)(v) = \log ( \vert f_n \vert e^{vn}) = u= \NP(f)^*(v).$

Therefore, for any $v \in \left] -\infty,v_{\infty} \right[,$ with $v$ not a slope of $\NP(f),$ we get $\Lambda (f)(v) =\NP(f)^*(v).$
Since $\Lambda (f)$ is increasing, $\NP(f)^*$ is continuous (since it is convex), and $\NP(f)$ has only finitely many slopes smaller than $v_\infty-\epsilon$, for any $\epsilon >0$, we deduce that both functions coincide on $v \in \left] -\infty,v_{\infty} \right[,$
\end{proof}

@@

The following proposition will reveal crucial in order to minorate or majorate $\NP(f),$  $\NP(f)^*$ or $\Lambda(f),$ and is classical in convex analysis.

@@

If we denote by $\NP((M_{n+1})_{n \geq 1})$ the lower convex hull of the $(n, -\log M_{n+1})$ for $n \geq 1$, then $\delta$ such that $-\log \varepsilon=\NP((M_n)_{n \geq 2})^*(r)$ is enough to define first order lattices as in lemma  \ref{lem:main}.
\end{prop}
\begin{proof}
Same idea as proposition \ref{prop : maj Lambda}. One can define $g_n$ with $g_n (x) =\frac{f_n(x)}{\vert x \vert}$ and $g_n(0)=0$ and work with $\Lambda ( \sum g_n).$
\end{proof}

@@

\begin{prop} \label{prop : locanalyticfinitedim}
If $f$ is a locally analytic map $K^n \rightarrow K^m$, with $K$ a $p$-adic field, if $C,$ $\rho$ and $\varepsilon$ are as above. It is enough to work with the Newton polygon $\NP_F$ of $F(t) = \sum_{k\geq 0}^{+\infty} \interleave f^{(k+1)}(0) \interleave  t^k \in \mathbb{R}[[t]]$ : let $r$ be such that $\NP_F^* \left( \log_p r - \frac{1}{p-1} \right) \leq \log_p \varepsilon.$ Then $r$ is enough for lemma \ref{lem:main}. 
\end{prop}
\begin{proof}
Same as before, with $L_n=\frac{f^{(n)}(0)}{n!}$ and Legendre formula. The result follows \textit{mutatis mutandis}.
\end{proof}

@@
